MSAD_PCA_Walkthrough

### 2017-07-13 LaTeX_Vignette/MS-MVSPC_DissertationVignette.Rnw
Edited section 5.2

\subsection{Why use \texttt{mvMonitoring}?}
The \texttt{mvMonitoring} package can be used to detect outliers in a correlated multivariate process with non-linear, non-stationary, or autocorrelated feature behavior. These process outliers are often indicative of system fault. The common  and naive approach to multivariate process monitoring is to use expert opinion to identify a few important features to monitor visually, and raise an alarm if these features travel outside pre-defined normal operating boundaries.

However, this split univariate approach fails to account for the correlated nature of the process features, so some engineers have taken to monitoring the system as a single correlated multivariate process rather than a collection of independent univariate processes. In the literature review of the motivating papers, the authors cite how this approach has benefited the science of process monitoring as a whole. Unfortunately, this approach has its own shortcomings -- namely, monitoring a multivariate process in its original feature space can lead to exorbitant computational costs. To this end, \emph{principal components analysis} (PCA) and its many modifications have been employed to reduce the computation necessary to monitor a multivariate process as a whole. This package is an implementation of one such PCA modification.

\subsection{Multi- or Single-State AD-PCA}
AD-PCA is \emph{adaptive-dynamic} PCA, which accounts for non-linearity, non-stationarity, and autocorrelation in the multivariate process, and this modification is thoroughly discussed in \cite{kazor_comparison_2016} as well as in the previous chapter. As an additional component to this model, we now consider that the process to monitor may have multiple known system states. Thus, \emph{multi-state} monitoring is a modification to PCA which accounts for multiple process states and models them separately. States can be any mutually exclusive blocking factor, and states do not necessarily follow a strict order. Then, Multi-State AD-PCA (MSAD-PCA) allows process engineers to account for a few very different process states. Specifically, this modification should be used when features under different process states have different means, correlations, variances, or some combination of these three.

>>>>

\subsection{Why use \texttt{mvMonitoring}?}
The \texttt{mvMonitoring} package can be used to detect outliers in a correlated, non-Gaussian multivariate process with non-linear, non-stationary, or autocorrelated feature behavior. These process outliers are often indicative of system fault. The naive (and unfortunately common) approach to multivariate process monitoring is to use expert opinion to identify a few important features to monitor visually, and raise an alarm if these features travel outside pre-defined normal operating boundaries.

However, this split univariate approach fails to account for the correlated nature of the process features, so some engineers have taken to monitoring the system as a single correlated multivariate process rather than a collection of independent univariate processes. In the literature review of the motivating papers, the authors cite how this approach has benefited the science of process monitoring as a whole. Unfortunately, this approach has its own shortcomings. One such complication is that monitoring a multivariate process in its original feature space can lead to exorbitant computational costs. Additionally, correlated multivariate process over many features may be noisy and often retain redundant information. To combat these issues, \emph{principal components analysis} (PCA) and its many modifications have been employed to assist in monitoring multivariate processes as a whole. This package is an implementation of one such PCA modification.

\subsection{Multi- or Single-State AD-PCA}
\emph{Adaptive-Dynamic} PCA (AD-PCA), thoroughly discussed in \cite{kazor_comparison_2016} as well as in the previous chapter, accounts for non-linearity, non-stationarity, and autocorrelation in non-Gaussian multivariate processes. As an additional layer of complexity within this model, consider such a process with multiple known system states. \emph{Multi-State} monitoring is thus a modification to PCA which accounts for multiple process states and models them separately. States can be any mutually exclusive blocking factor, and states do not necessarily follow a strict order. Then, Multi-State AD-PCA (MSAD-PCA) allows process engineers to account for distinct process states. Specifically, this modification should be used when features under different process states have different means, correlations, variances, or some combination of these three.

### 2017-07-13 LaTeX_Vignette/MS-MVSPC_DissertationVignette.tex
Edited section 5.2

##   Data: num [1:4142, 1:5] 0.042 0.492 1.208 1.139 0.742 ...

@@

##   .. ..$ SPE_threshold   : Named num 4.39

@@

##   .. ..$ T2_threshold    : Named num 61.4

@@

##   .. ..$ projectionMatrix: num [1:6, 1:2] 0.424 -0.381 0.437 0.432 -0.313 ...
##   .. ..$ LambdaInv       : num [1:2, 1:2] 0.206 0 0 1.565
##   .. ..$ muTrain         : Named num [1:6] 1.38 -2.09 2.97 1.37 -2.05 ...

@@

##   .. ..$ RootPrecisTrain : num [1:6, 1:6] 2.58 0 0 0 0 ...

@@

##   .. ..$ SPE_threshold   : Named num 1.79

@@

##   .. ..$ T2_threshold    : Named num 76.6

@@

##   .. ..$ projectionMatrix: num [1:6, 1:3] -0.4944 -0.0882 0.4873 -0.5234 -0.1921 ...
##   .. ..$ LambdaInv       : num [1:3, 1:3] 0.283 0 0 0 0.727 ...
##   .. ..$ muTrain         : Named num [1:6] 3.019 -0.561 -4.533 2.991 -0.586 ...

@@

##   .. ..$ RootPrecisTrain : num [1:6, 1:6] 1.07 0 0 0 0 ...

@@

##   .. ..$ SPE_threshold   : Named num 5.58

@@

##   .. ..$ projectionMatrix: num [1:6, 1:2] 0.394 0.448 0.452 -0.167 0.464 ...
##   .. ..$ LambdaInv       : num [1:2, 1:2] 0.277 0 0 0.498
##   .. ..$ muTrain         : Named num [1:6] 0.528 0.273 1.441 0.57 0.259 ...

@@

##   .. ..$ RootPrecisTrain : num [1:6, 1:6] 17.2 0 0 0 0 ...

@@

\subsubsection{Adding Lagged Features}
First, concatenate the last given observation from the training set as ``row 0'' of the test data set. This will enable \texttt{mspMonitor} to include lag-1 features. Similarly, one would include the last $k$ observations of the training set should the process dictate the need for any lag-$k$ features. Because the Fault Start Index was set to 8500, this testing window will show the change point between observations generated under normal conditions and those generated under a fault state.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{test1A_xts} \hlkwb{<-} \hlstd{fault1A_xts[}\hlnum{8460}\hlopt{:}\hlnum{8520}\hlstd{,} \hlopt{-}\hlnum{1}\hlstd{]}
\hlstd{lagTest1A_xts} \hlkwb{<-} \hlkwd{lag.xts}\hlstd{(test1A_xts,} \hlnum{0}\hlopt{:}\hlnum{1}\hlstd{)}
\hlstd{lagTest1A_xts} \hlkwb{<-} \hlkwd{cbind}\hlstd{(fault1A_xts[}\hlnum{8461}\hlopt{:}\hlnum{8520}\hlstd{,}\hlnum{1}\hlstd{],}
                       \hlstd{lagTest1A_xts[}\hlopt{-}\hlnum{1}\hlstd{,])}
\hlkwd{head}\hlstd{(lagTest1A_xts)}
\end{alltt}
\begin{verbatim}
##                     state         x         y         z       x.1
## 2015-05-22 07:00:00     1 0.6516822 -1.487152 0.8868843 0.3772442
## 2015-05-22 07:01:00     1 0.6959653 -1.524802 0.7687056 0.6516822
## 2015-05-22 07:02:00     1 0.4693409 -1.202725 0.7970705 0.6959653
## 2015-05-22 07:03:00     1 0.5744935 -1.239971 0.6663379 0.4693409
## 2015-05-22 07:04:00     1 0.5872442 -1.238700 0.8654388 0.5744935
## 2015-05-22 07:05:00     1 0.6820071 -1.452199 0.9107379 0.5872442
##                            y.1       z.1
## 2015-05-22 07:00:00  0.1208201 0.5190793
## 2015-05-22 07:01:00 -1.4871516 0.8868843
## 2015-05-22 07:02:00 -1.5248019 0.7687056
## 2015-05-22 07:03:00 -1.2027252 0.7970705
## 2015-05-22 07:04:00 -1.2399706 0.6663379
## 2015-05-22 07:05:00 -1.2386996 0.8654388
\end{verbatim}
\end{kframe}
\end{knitrout}

\subsubsection{Monitoring the Test Data}
With the lagged test observations in the working environment, the \texttt{mspMonitor} function can be applied. This function (similarly to \texttt{mspTrain}) takes in the label information as a separate argument from the input data. Further, the \texttt{mspMonitor} function takes in the \texttt{TrainingSpecs} object returned in the results list from \texttt{mspTrain}. Notice that the first six rows of the monitor matrix are the exact same as the first six rows of the lagged test matrix, except that the rows of the monitor matrix have the monitoring statistic and corresponding indicator columns appended.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{monitor1A_xts} \hlkwb{<-} \hlkwd{mspMonitor}\hlstd{(}\hlkwc{observations} \hlstd{= lagTest1A_xts[,}\hlopt{-}\hlnum{1}\hlstd{],}
                            \hlkwc{labelVector} \hlstd{= lagTest1A_xts[,}\hlnum{1}\hlstd{],}
                            \hlkwc{trainingSummary} \hlstd{= train1A_ls}\hlopt{$}\hlstd{TrainingSpecs)}
\hlkwd{head}\hlstd{(monitor1A_xts)}
\end{alltt}
\begin{verbatim}
##                             x         y         z       x.1        y.1
## 2015-05-22 07:00:00 0.6516822 -1.487152 0.8868843 0.3772442  0.1208201
## 2015-05-22 07:01:00 0.6959653 -1.524802 0.7687056 0.6516822 -1.4871516
## 2015-05-22 07:02:00 0.4693409 -1.202725 0.7970705 0.6959653 -1.5248019
## 2015-05-22 07:03:00 0.5744935 -1.239971 0.6663379 0.4693409 -1.2027252
## 2015-05-22 07:04:00 0.5872442 -1.238700 0.8654388 0.5744935 -1.2399706
## 2015-05-22 07:05:00 0.6820071 -1.452199 0.9107379 0.5872442 -1.2386996
##                           z.1       SPE SPE_Flag        T2 T2_Flag Alarm
## 2015-05-22 07:00:00 0.5190793 0.5746873        0 28.479963       0    NA
## 2015-05-22 07:01:00 0.8868843 0.3065558        0  4.594196       0    NA
## 2015-05-22 07:02:00 0.7687056 2.2496108        0  6.313058       0    NA
## 2015-05-22 07:03:00 0.7970705 1.4836139        0  7.071992       0    NA
## 2015-05-22 07:04:00 0.6663379 1.6703320        0  6.605812       0    NA
## 2015-05-22 07:05:00 0.8654388 0.4398132        0  5.638591       0    NA
\end{verbatim}
\end{kframe}
\end{knitrout}

>>>>

##   Data: num [1:4142, 1:5] 2.084 0.268 0.582 0.694 0.956 ...

@@

##   .. ..$ SPE_threshold   : Named num 4.66

@@

##   .. ..$ T2_threshold    : Named num 58.1

@@

##   .. ..$ projectionMatrix: num [1:6, 1:2] 0.421 -0.386 0.434 0.429 -0.325 ...
##   .. ..$ LambdaInv       : num [1:2, 1:2] 0.202 0 0 1.685
##   .. ..$ muTrain         : Named num [1:6] 1.38 -2.08 2.95 1.36 -2.04 ...

@@

##   .. ..$ RootPrecisTrain : num [1:6, 1:6] 2.47 0 0 0 0 ...

@@

##   .. ..$ SPE_threshold   : Named num 1.06

@@

##   .. ..$ T2_threshold    : Named num 77.4

@@

##   .. ..$ projectionMatrix: num [1:6, 1:3] -0.50267 -0.00244 0.49616 -0.52609 -0.14826 ...
##   .. ..$ LambdaInv       : num [1:3, 1:3] 0.285 0 0 0 0.706 ...
##   .. ..$ muTrain         : Named num [1:6] 2.958 -0.564 -4.481 2.931 -0.59 ...

@@

##   .. ..$ RootPrecisTrain : num [1:6, 1:6] 1.04 0 0 0 0 ...

@@

##   .. ..$ SPE_threshold   : Named num 6.33

@@

##   .. ..$ projectionMatrix: num [1:6, 1:2] -0.403 -0.456 -0.459 0.135 -0.458 ...
##   .. ..$ LambdaInv       : num [1:2, 1:2] 0.273 0 0 0.505
##   .. ..$ muTrain         : Named num [1:6] 0.523 0.269 1.409 0.565 0.255 ...

@@

##   .. ..$ RootPrecisTrain : num [1:6, 1:6] 15.4 0 0 0 0 ...

@@

\subsubsection{Adding Lagged Features}
First, concatenate the last given observation from the training set as ``row 0'' of the test data set. This will enable \texttt{mspMonitor} to include lag-1 features. Similarly, one would include the last $k$ observations of the training set should the process dictate the need for any lag-$k$ features. Because the Fault Start Index was set to 8500, this testing window will show the change point between observations generated under normal conditions and those generated under a fault state.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{test1A_xts} \hlkwb{<-} \hlstd{fault1A_xts[}\hlnum{8460}\hlopt{:}\hlnum{8520}\hlstd{,} \hlopt{-}\hlnum{1}\hlstd{]}
\hlstd{lagTest1A_xts} \hlkwb{<-} \hlkwd{lag.xts}\hlstd{(test1A_xts,} \hlnum{0}\hlopt{:}\hlnum{1}\hlstd{)}
\hlstd{lagTest1A_xts} \hlkwb{<-} \hlkwd{cbind}\hlstd{(fault1A_xts[}\hlnum{8461}\hlopt{:}\hlnum{8520}\hlstd{,}\hlnum{1}\hlstd{],}
                       \hlstd{lagTest1A_xts[}\hlopt{-}\hlnum{1}\hlstd{,])}
\hlkwd{head}\hlstd{(lagTest1A_xts)}
\end{alltt}
\begin{verbatim}
##                     state         x         y         z       x.1
## 2015-05-22 07:00:00     1 0.6030535 -1.566528 0.8662376 0.3665282
## 2015-05-22 07:01:00     1 0.4416746 -1.406672 0.3811997 0.6030535
## 2015-05-22 07:02:00     1 0.5408433 -1.304789 0.4314055 0.4416746
## 2015-05-22 07:03:00     1 0.5082937 -1.391546 0.8054482 0.5408433
## 2015-05-22 07:04:00     1 0.4508361 -1.312723 0.8291703 0.5082937
## 2015-05-22 07:05:00     1 0.7068563 -1.465259 0.7728310 0.4508361
##                             y.1       z.1
## 2015-05-22 07:00:00  0.08751379 0.3336879
## 2015-05-22 07:01:00 -1.56652763 0.8662376
## 2015-05-22 07:02:00 -1.40667237 0.3811997
## 2015-05-22 07:03:00 -1.30478897 0.4314055
## 2015-05-22 07:04:00 -1.39154587 0.8054482
## 2015-05-22 07:05:00 -1.31272320 0.8291703
\end{verbatim}
\end{kframe}
\end{knitrout}

\subsubsection{Monitoring the Test Data}
With the lagged test observations in the working environment, the \texttt{mspMonitor} function can be applied. This function (similarly to \texttt{mspTrain}) takes in the label information as a separate argument from the input data. Further, the \texttt{mspMonitor} function takes in the \texttt{TrainingSpecs} object returned in the results list from \texttt{mspTrain}. Notice that the first six rows of the monitor matrix are the exact same as the first six rows of the lagged test matrix, except that the rows of the monitor matrix have the monitoring statistic and corresponding indicator columns appended.
\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlstd{monitor1A_xts} \hlkwb{<-} \hlkwd{mspMonitor}\hlstd{(}\hlkwc{observations} \hlstd{= lagTest1A_xts[,}\hlopt{-}\hlnum{1}\hlstd{],}
                            \hlkwc{labelVector} \hlstd{= lagTest1A_xts[,}\hlnum{1}\hlstd{],}
                            \hlkwc{trainingSummary} \hlstd{= train1A_ls}\hlopt{$}\hlstd{TrainingSpecs)}
\hlkwd{head}\hlstd{(monitor1A_xts)}
\end{alltt}
\begin{verbatim}
##                             x         y         z       x.1         y.1
## 2015-05-22 07:00:00 0.6030535 -1.566528 0.8662376 0.3665282  0.08751379
## 2015-05-22 07:01:00 0.4416746 -1.406672 0.3811997 0.6030535 -1.56652763
## 2015-05-22 07:02:00 0.5408433 -1.304789 0.4314055 0.4416746 -1.40667237
## 2015-05-22 07:03:00 0.5082937 -1.391546 0.8054482 0.5408433 -1.30478897
## 2015-05-22 07:04:00 0.4508361 -1.312723 0.8291703 0.5082937 -1.39154587
## 2015-05-22 07:05:00 0.7068563 -1.465259 0.7728310 0.4508361 -1.31272320
##                           z.1       SPE SPE_Flag        T2 T2_Flag Alarm
## 2015-05-22 07:00:00 0.3336879 0.7953977        0 24.566074       0    NA
## 2015-05-22 07:01:00 0.8662376 0.3442694        0  5.731867       0    NA
## 2015-05-22 07:02:00 0.3811997 0.4382832        0  6.232419       0    NA
## 2015-05-22 07:03:00 0.4314055 0.2334754        0  5.558003       0    NA
## 2015-05-22 07:04:00 0.8054482 0.5248950        0  5.476091       0    NA
## 2015-05-22 07:05:00 0.8291703 0.1880696        0  4.896475       0    NA
\end{verbatim}
\end{kframe}
\end{knitrout}
